Exacerbated by the Pandemic: Analyzing the Impact of COVID-19 on Mortality Among Toronto's Homeless by Gender Disparities

### 2024-03-17 paper/paper.qmd
paper.qmd

This study aims to explore the impact of the COVID-19 on the difference of the death rate between female and male. Employing the “Difference in Differences” approach, I will analyze homeless death data in Toronto, spanning from January 2017 to June 2023. We assume two time periods: 2017 January to 2019 December is before covid; 2020 January to 2023 June is in the covid. Death count is the estimand. We will embark on an analysis to validate the hypothesis that the COVID-19 pandemic has led to an increase in the death count among the homeless population. We will present this analysis in two distinct parts: one without controlling for gender, to capture the overall trend, and another with gender control, to discern any differential impact between male and female homeless individuals. Firstly, we will inspect the aggregated data, where we anticipate an observable escalation in death counts following the outbreak, regardless of gender or age group differences. Subsequently, we will refine our analysis with a gender-based perspective. Here, we aim to unravel the nuances of how the pandemic may have disparately affected homeless males and females. This part of the analysis seeks to highlight the intersection of gender with the broader health crisis. In conjunction with these visual comparisons, we will also delve into a more granular analysis by presenting the results of a regression model. This model will quantify the impact of COVID-19 on death counts, first without the inclusion of gender, and then with gender considered. Through this comparative approach, we aim to discern whether the observed increase in death counts post-pandemic holds when we account for gender, as well as other control variables such as the cause of death and age. The conclusions drawn will be critical, not just in understanding the impact of the pandemic but also in informing potential interventions and policy decisions aimed at mitigating the risks faced by homeless populations in such health crises.

>>>>

This study aims to explore the impact of the COVID-19 on the difference of the death rate between female and male. Employing the “Difference in Differences” approach, I will analyze homeless death data in Toronto, spanning from January 2017 to June 2023. We assume two time periods: 2017 January to 2019 December is before covid; 2020 January to 2023 June is in the covid. Death count is the estimand. We will embark on an analysis to validate the hypothesis that the COVID-19 pandemic has led to an increase in the death count among the homeless population. We will present this analysis in two distinct parts: one without controlling for gender, to capture the overall trend (Figure 6), and another with gender control, to discern any differential impact between male and female homeless individuals (Figure 7). Firstly, we will inspect the aggregated data, where we anticipate an observable escalation in death counts following the outbreak, regardless of gender or age group differences. Subsequently, we will refine our analysis with a gender-based perspective. Here, we aim to unravel the nuances of how the pandemic may have disparately affected homeless males and females. This part of the analysis seeks to highlight the intersection of gender with the broader health crisis. In conjunction with these visual comparisons, we will also delve into a more granular analysis by presenting the results of a regression model. This model will quantify the impact of COVID-19 on death counts, first without the inclusion of gender, and then with gender considered. Through this comparative approach, we aim to discern whether the observed increase in death counts post-pandemic holds when we account for gender, as well as other control variables such as the cause of death and age (Table 2). The conclusions drawn will be critical, not just in understanding the impact of the pandemic but also in informing potential interventions and policy decisions aimed at mitigating the risks faced by homeless populations in such health crises.
